---
title: "Problem Set 3"
author: "Ghina Al Shdaifat - gha2009 - 003 - Collaborated: Petra Ivanovic(pi2018)&Kendall Brogle (klb9313)"
date: "Due Dec 1, 2023"
output:
  pdf_document: 
    latex_engine: lualatex
header-includes: \usepackage{tikz}
---

This homework must be turned in on Brightspace by Dec. 1, 2023. It must be your own work, and your own work only -- you must not copy anyone's work, or allow anyone to copy yours. This extends to writing code. You may consult with others, but when you write up, you must do so alone.

Your homework submission must be written and submitted using Rmarkdown. No handwritten solutions will be accepted. \textbf{No zip files will be accepted.} \textbf{Make sure we can read each line of code in the pdf document.} You should submit the following:

1. A compiled PDF file named yourNetID\_solutions.pdf containing your solutions to the problems.

2. A .Rmd file containing the code and text used to produce your compiled pdf named yourNetID\_solutions.Rmd.


Note that math can be typeset in Rmarkdown in the same way as Latex. Please make sure your answers are clearly structured in the Rmarkdown file:

1. Label each question part

2. Do not include written answers as code comments.

3. The code used to obtain the answer for each question part should accompany the written answer. Comment your code!


\newpage 


## Question 1 (Total: 100)
Does US military assistance strengthen or further weaken fragile and conflict-affected foreign governments? Aid may bolster state capacity and suppress violence from nonstate actors such as paramilitary groups. On the other hand, aid may be diverted to those same violent groups. To answer the question, Dube and Naidu (2015)(https://www.journals.uchicago.edu/doi/10.1086/679021?mobileUi=0) leverage changes in the allocation of US military aid to Colombian military bases. They test whether Colombian municipailites in which military bases are located have more or less paramilitary violence when the level of U.S. miliary aid increases, relative to Colombian municipalities in which miliary bases are not located. 

For this problem, you will need the `bases_replication_file.dta' file. The variables you will need are:

- parattq - DV here is paramilitary attacks
- bases6 - indicator variable whether or not there is a base in the municipality
- lrmilnar col - (logged) U.S. military and narcotics aid to Colombia
- bases6xlrmilnar col - the treatment i.e., the interaction between the level of U.S. military and narcotics aid and whether or not there is a base in the municipality
- lnnewpop - is log of population

### Part a (60 points)

The treatment in this case is a continuous 'intensity' variable that changes over time. The authors use the interaction between the level of U.S. military and narcotics aid and whether a base exists in a municipality. How many units are in the 'control' group (no bases)? Does the bases variable change over time or is it a unit-constant factor? How about the logged military aid variable, does it change across units for a given year? What do the authors seem to be assuming about how military aid is allocated?

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(haven)
library(estimatr) # for lm with robust se : ?lm_robust()

# Load bases data
bases <- haven::read_dta("/Users/ghinaalshdaifat/Desktop/bases_replication_final.dta")
# Filtering out variables # Selecting specific columns
bases <- bases %>%
select(municipality, year,paratt, bases6, lrmilnar_col, bases6xlrmilnar_col, 
       lnnewpop)
bases
```
```{r}
# How many observations are in the "no bases group"
noBases_count <- bases %>% filter(bases6 == 0) %>% nrow()
noBases_count
```
```{r}
# Count municipalities with zero
municipalitiesWzero <- bases %>%
  group_by(municipality) %>% # Group by municipality 
  summarise(all_zero = all(bases6 == 0)) %>% # Filter with base equal to 0
  filter(all_zero) %>% 
  nrow() 
# Print the number of municipalities with zero
print(municipalitiesWzero)
```
- 16272 values are in the “no bases group,” indicating that 16272 of our units 
do not have bases. 
- Filtering out the distinct municipalities indicates a total of 904 
municipalities that have no bases. 
```{r}
## How about each of them?
change <- bases %>% 
  group_by(municipality) %>% 
  summarise(unique_bases_values = n_distinct(bases6))
change
```
```{r}
countNon1Values <- change %>% filter(unique_bases_values 
                                                        != 1) %>% nrow()
# Number of municipalities with a base of value of > 1
countNon1Values
```
There is a single unique base value for each of our municipalities, indicating
no change as time passes. With a count of 0, we can additionally illustrate that
there are no municipalities with any number of unique bases values but 1. 
Therefore, we can conclude that the bases are unit-constant for each individual
municipality. In other words, status of military bases do not change as time
passes by for any of the municipalities. The bases variable is constant for all
municipalities for the time period presented in the data. 

```{r}
# Examining changes in logged military aid across units
aidVar <- bases %>%
  group_by(year, municipality) %>% summarise(unique_aid_values = 
                                             n_distinct(lrmilnar_col))
aidVar
```
```{r}
# Number of municipalities with multiple aid values
nonUniqueAidCount <- aidVar %>% filter(unique_aid_values != 1) %>% nrow()
print(paste(nonUniqueAidCount))
```
The logged value of military aid remains constant across municipalities. 
This is evidenced by the "non_unique_aid_count" being zero, suggesting that for 
each grouping by year and municipality, there is no variation in military aid – 
all municipalities receive an identical amount of aid within a given year. 
Nonetheless, this uniformity in aid distribution doesn't imply that the amount 
of aid received was consistent across different years. Further examination is 
required to understand how aid distribution varied year over year.

```{r}
aidVarAll <- bases %>% group_by(municipality) %>%
  summarise(unique_aid_values = n_distinct(lrmilnar_col))

aidVarAll
```
```{r}
# Number of municipalities that have more or less than 18 aid value
non18 <- aidVarAll %>% filter(unique_aid_values != 18) %>% nrow()
non18
```
```{r}
# Aid for municipality for the year 5113
municipality5113 <- bases %>% filter(municipality == 5113) %>% arrange(year)
municipality5113
```
Without considering the year, each municipality experiences 18 distinct 
instances of changes in military aid. This pattern suggests that the 
fluctuations in aid likely occur at the beginning or end of each year, given the 
observation that grouping the data by year results in a single unique value of 
aid per municipality, implying stability within the same year. Additionally, the 
second table provides a detailed view for a specific municipality, number 5113, 
chosen for illustrative purposes. The data for this municipality aligns with the 
previous observation, displaying 18 separate entries, each corresponding to a 
different year, and showing variations in the logged military aid value, while 
the "bases" value remains unchanged. Overall, the analysis confirms that the 
quantity of military aid allocated to each municipality remains consistent 
throughout any given year but varies from year to year.

```{r}

## How many municipalities do we have
numbMunicipalities <- bases %>% distinct(municipality) %>% nrow()
numbMunicipalities
```
The dataset includes information on 936 distinct municipalities. It was 
determined that 904 of these municipalities do not have any military bases,
hence, there are bases present in 32 of the municipalities.

The researchers appear to be working under the presumption that the distribution 
of military aid to each municipality remains the same throughout a given year,
but shifts from one year to the next. This inference is drawn from the 
identified trend where each municipality is allocated a distinct value of 
military aid that does not vary within that particular year, although it does so 
across different years. This indicates there is a notion that the military aid 
is evenly distributed (constant for each unit) during a single year, followed by 
adjustments in subsequent years. Such an assumption of annual uniformity 
suggests that the process of allocating aid might not take into account the 
varying needs that could emerge within municipalities over the course of the 
same year.

### Part b (20 points)
The authors use a common empirical strategy called two-way fixed effects to estimate the average treatment effect of military aid. The model they estimate includes fixed effects for both time periods and units (and includes logged population as an additional covariate):

$$ Y_{it} = \gamma_t + \alpha_i + \tau D_{it} + \beta X_{it} + \epsilon_{it}$$ 
What assumptions are the authors making in order to identify the treatment effect of military aid?

The authors' methodology includes the use of fixed effects in their statistical 
model, suggesting they believe that by incorporating these effects, they are 
adjusting for all immutable characteristics of the municipalities that might 
also impact paramilitary violence. These characteristics are constant within 
each municipality over time, which helps separate the impact of U.S. aid from 
other variables.

Regarding parallel trends, the assumption is that in the absence of changes in 
U.S. aid, the municipalities with military bases would exhibit trends in 
paramilitary violence similar to those without bases. If this assumption holds, 
it implies that observed differences in trends post-aid allocation are 
attributable to the aid.

The authors also presume that the error term (\( \epsilon_{it} \)) captures all 
the unobserved factors that could influence paramilitary attacks, such as 
economic conditions. This presumes these unobserved factors are not correlated 
with the military aid, suggesting that the results should be free from bias 
related to these unmeasured variables.

Furthermore, the study assumes the principle of SUTVA (Stable Unit Treatment 
Value Assumption) and independence of treatment assignment. This means they 
suppose that an increase in military aid in one municipality doesn't affect 
paramilitary violence in another, indicating no spillover effects. It also 
implies that the allocation of military aid is not based on any characteristics 
of the municipalities that simultaneously affect their levels of paramilitary 
violence, such as a heightened need for aid. Essentially, they are considering 
the allocation of military bases and the consequent aid to be independent of the 
outcomes they are measuring.

They also assumed controlled treatment effect - (aid) is constant between units 
and time.

### Part c (20 points)

Using the two-way fixed effects estimator, estimate the effect of U.S. military and narcotics aid on the number of paramilitary attacks, including log of population as a covariate. The two sets of fixed effects are for municipality (municipality) and year (year). Cluster your standard errors at the unit level (see the cluster argument in lm_robust. Report a 95% confidence interval for your estimate and interpret your results.

```{r}
#?lm_robust (set se_type to "CR0")
# Fit Regression using lm_robust 

# Fit the linear model with robust standard errors
fitted_model <- lm_robust(
  paratt ~ bases6 + lrmilnar_col + bases6xlrmilnar_col + lnnewpop + 
    factor(municipality) + factor(year),
  data = bases,
  clusters = municipality,
  se_type = "CR0"
)

# Summarize the fitted model
summary_fitted <- summary(fitted_model)

# Retrieve the model's coefficients
coefficients_fitted <- summary_fitted$coefficients

# Filter out factor levels from the coefficients to shorten the output
main_coefficients <- coefficients_fitted[!grepl("factor", 
                                              rownames(coefficients_fitted)), ]

# Output the main coefficients
print(main_coefficients)

# Report the R-squared values
cat("Computed R-squared:", summary_fitted$r.squared, "\n")
cat("Computed Adjusted R-squared:", summary_fitted$adj.r.squared, "\n")


```
The coefficient for the variable indicating the presence of military bases 
(bases6) stands at roughly 2.801, with a standard error near 0.270. This 
reflects a positive association between military bases and the dependent
variable, which in this context is the frequency of paramilitary attacks. 
The statistical significance of this variable is confirmed by a t-value of 
10.367 and a p-value well below 0.001.

Regarding the variable representing logged military aid (lrmilnar_col), the 
coefficient is near -0.115 with a standard error around 0.057, pointing to a 
negative correlation with the dependent variable. The reduction in paramilitary 
attacks with increased military aid is statistically significant, as the p-value 
falls below 0.05.

For the interaction term between the presence of bases and logged military aid
(bases6xlrmilnar_col), the coefficient is close to 0.150 with a standard error 
of 0.060. This positive coefficient infers that the impact of military aid on 
paramilitary attacks is contingent upon whether a municipality has a military 
base, with the effect being statistically significant (p-value under 0.05).

The logged population size (lnnewpop) yields a coefficient of around 0.118 and a 
standard error of 0.045, suggesting that an increase in population size 
correlates with a rise in paramilitary attacks, with statistical significance 
indicated by a p-value less than 0.01.

As for the model's explanatory power, the multiple R-squared value is 
approximately 0.221, meaning the model accounts for about 22.1% of the variation 
in the dependent variable. After adjusting for the number of predictors, the 
adjusted R-squared value is about 0.174, signifying that roughly 17.4% of the 
variation is explained.

Each coefficient is accompanied by a 95% confidence interval, 
delineating the range within which we can be 95% certain the actual parameter 
value for the population is found. Additionally, degrees of freedom, listed as 
935 for each coefficient, are crucial for the computation of both the t-value 
and the subsequent p-value.

Overall, the analysis infers that military bases correlate with a rise in 
paramilitary attacks, that increased military aid is linked to a decrease in 
such attacks, but this link is altered when a base is present. Furthermore, the 
model illustrates a correlation between higher population sizes and an uptick in 
paramilitary attacks. The R-squared values indicate that while the model sheds 
light on some of the variation in the dependent variable, there remains a 
notable portion that is unaccounted for.

