---
title: "Problem Set 2"
author: "Ghina Al Shdaifat - gha2009 - 003 "
date: "Due Nov 10, 2023"
output:
  pdf_document: 
    latex_engine: lualatex
header-includes: \usepackage{tikz}
---

This homework must be turned in on Brightspace by Nov. 10, 2023. It must be your own work, and your own work only -- you must not copy anyone's work, or allow anyone to copy yours. This extends to writing code. You may consult with others, but when you write up, you must do so alone.

Your homework submission must be written and submitted using Rmarkdown. No handwritten solutions will be accepted. \textbf{No zip files will be accepted.} \textbf{Make sure we can read each line of code in the pdf document.} You should submit the following:

1. A compiled PDF file named yourNetID\_solutions.pdf containing your solutions to the problems.

2. A .Rmd file containing the code and text used to produce your compiled pdf named yourNetID\_solutions.Rmd.


Note that math can be typeset in Rmarkdown in the same way as Latex. Please make sure your answers are clearly structured in the Rmarkdown file:

1. Label each question part

2. Do not include written answers as code comments.

3. The code used to obtain the answer for each question part should accompany the written answer. Comment your code!


\newpage 


## Question 1 (Total: 50)
In new democracies and post-conflict settings, Truth and Reconciliation Commissions (TRCs) are often
tasked with investigating and reporting about wrongdoing in previous governments. Depending on the context, institutions such as TRCs are expected to reduce hostilities (e.g. racial hostilities) and promote peace.

In 1995, South Africa’s new government formed a national TRC in the aftermath of apartheid. [Gibson 2004] uses survey data collected from 2000-2001 to examine whether this TRC promoted inter-racial reconciliation. The outcome of interest is respondent racial attitudes (as measured by the level of agreement with the prompt: "I find it difficult to understand the customs and ways of [the opposite racial group]".) The treatment is ``exposure to the TRC" as measured by the individual’s level of self-reported knowledge about the TRC.

You will need to use the trc_data.dta file for this question. The relevant variables are:

- RUSTAND - Outcome: respondent’s racial attitudes (higher values indicate greater agreement)
- TRCKNOW - Treatment dummy (1 = if knows about the TRC, 0 = otherwise)
- age - Respondent age (in 2001)
- female - Respondent gender
- wealth - Measure of wealth constructed based on asset ownership (assets are fridge, floor polisher,
    vacuum cleaner, microwave oven, hi-fi, washing machine, telephone, TV, car)
- religiosity - Self-reported religiosity (7 point scale)
- ethsalience - Self-reported ethnic identification (4 point scale)
- rcblack - Respondent is black
- rcwhite - Respondent is white
- rccol - Respondent is coloured (distinct multiracial ethnic group)
- EDUC - Level of education (9 point scale)

### Part a (15 points)

Estimate the average treatment effect of TRC exposure on respondents’ racial attitudes under the assumption that TRC exposure is ignorable. Report a 95% confidence interval for your estimate and interpret your results. (Use robust standard errors throughout.)


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(haven)
library(estimatr)

# Load in the TRC data (it's a STATA .dta so we use the haven package)
TRC_data <- haven::read_dta("trc_data.dta")

# Estimate the average treatment effect of TRC exposure
# HC1 -> to ensure standard error is robust 
model <- lm_robust(RUSTAND ~ TRCKNOW, data = TRC_data, se_type = "HC1")

# Output the results with a 95% confidence interval
tidy(model, conf.int = TRUE)

```
Interpretation: 

- The coefficient for the treatment dummy ('TRCKNOW) is -0.2177, with a standard 
error of 0.04433. In this case, the negative sign indicates that as knowledge 
about the TRC increases, the level of agreement decreases, indicating an inverse
relationship between both the treatment and outcome and hence suggesting an 
improvement in racial attitudes.

- The corresponding p-value is approximately 0.0000009489, which is much lower 
than the alpha threshold of 0.05, indicating that this is a statistically 
significant result.

- The 95% confidence interval for the `TRCKNOW` coefficient ranges from -0.3047 
to -0.1308, which means we can say with 95% confidence that the true effect of 
TRC knowledge on racial attitudes lies within this interval. Since the interval 
does not include 0, this further validates that the effect is statistically 
significant.

Overall, the results indicate that there is a statistically significant 
association between knowledge of the TRC and racial attitudes, with greater 
knowledge correlating with more understanding towards customs and ways of the 
opposite racial group. In other words, TRC reduced respondent’s racial attitudes 
by 0.2177. 

### Part b (15 points)

Examine whether exposed and non-exposed respondents differ on the full set of observed covariates using a series of balance tests. Briefly discuss, in which ways do exposed and nonexposed respondents differ?


```{r}
balance_results <- data.frame(Covariate = character(),
                              Mean_Exposed = numeric(), 
                              Mean_Non_Exposed = numeric(), 
                              T_Value = numeric(), 
                              P_Value = numeric(),
                              stringsAsFactors = FALSE)

covariates <- c("age", "female", "wealth", "religiosity", "ethsalience", 
                "rcblack", "rcwhite", "rccol", "EDUC")

for (covariate in covariates){
  exposed_data <- TRC_data %>% filter(TRCKNOW == 1) %>% pull(!!sym(covariate)) 
  non_exposed_data <- TRC_data %>% filter(TRCKNOW == 0) %>% 
    pull(!!sym(covariate))
 
  # Mean 
  mean_exposed <- mean(exposed_data, na.rm = TRUE) 
  mean_non_exposed <- mean(non_exposed_data, na.rm = TRUE)

  # t-test
  test_result <- t.test(exposed_data, non_exposed_data, var.equal = TRUE)

# Append data frame with values
  balance_results <- balance_results %>% 
    add_row(Covariate = covariate,
            Mean_Exposed = mean_exposed, 
            Mean_Non_Exposed = mean_non_exposed, 
            T_Value = test_result$statistic, 
            P_Value = test_result$p.value)
}

# Print output
print(balance_results)

```
Interpretation: 

The analysis of the dataset reveals significant imbalances between the treated 
and control groups. Utilizing a p-value threshold of 0.05, notable differences 
are observed in several key attributes: age, female gender, wealth, rcblack, 
rccol, and EDUC. Considering that out of nine examined attributes, six exhibit 
statistically significant differences in their group averages, it is clear that 
the groups are imbalanced. 

Specifically, the age group exposed to the treatment is, on average, 
approximately a year and a half younger than their non-exposed counterparts,
indicated by an absolute t-value of 2.76. The table above also demonstrates that 
the exposed group contains approximately 10% more females, with a t-value of 
5.96, while simultaneously being wealthier. The proportion of black respondents 
is 55.18% in the exposed group and 51.30% in the non-exposed group. The positive 
t-value (2.187198) and p-value (0.02880017) indicate a significant difference;
on the other hand, white respondents have no significant difference with a 
p-value of 0.2894507 and a t-value of 1.059503. Generally, there is a lower
proportion of colored respondents in the exposed group and this difference
(11.05% of the exposed group and 15.74%) is significant with a p-value of
0.0001164664 and a negative t-value of -3.858219. Finally, the mean education 
level is higher for the exposed group (4.2919) compared to the non-exposed group 
(3.8465). The high positive t-value (11.015352) and very low p-value 
(1.007117e-27) suggest a significant difference, indicating higher education 
levels among the exposed group.

Overall, we can observe that there are indeed significant differences between 
the exposed and non-exposed respondents in regard to specific covariates. The 
groups that are not balanced on the other covariates could potentially influence 
the assumption of ignorability and hence lead to inaccuracies in identifying 
casual inferences that are drawn from the treatment effect analysis. Since there
are differences in statistical significance between both groups, we can assume
that the groups are not balanced. 


### Part c (10 points)

Now assume that TRC exposure is conditionally ignorable given the set of observed covariates:

1. Use a logistic regression model to estimate the propensity score for each observation. (For purposes of this question, do not include any interactions.)
2. With this model, construct inverse propensity of treatment weights (IPTW) for each observation using the unstabilized weights.
3. Use the propensity score to construct an IPW estimator and report the point estimate for the ATE.

Use the following covariates: age, female, wealth, religiosity, ethsalience, rcblack, rcwhite, rccol, EDUC


```{r}
# 1. Logistic regression model
propensity_model <- glm(TRCKNOW ~ age + female + wealth + religiosity + 
                        ethsalience + rcblack + rcwhite + rccol + EDUC, data = 
                          TRC_data, family = binomial(link = "logit"))
# Extract predicted propensity scores
TRC_data$propensity_score <- predict(propensity_model, type = "response")

# 2. IPTW using unstabalized weights
TRC_data$iptw <- ifelse(TRC_data$TRCKNOW == 1, 
                        1 / TRC_data$propensity_score, 
                        1 / (1 - TRC_data$propensity_score))

# 3. Calculate the IPTW ATE using weighted means
TRC_data <- TRC_data %>%
  mutate(weighted_outcome_treat = ifelse(TRCKNOW == 1, RUSTAND * iptw, 0),
         weighted_outcome_control = ifelse(TRCKNOW == 0, RUSTAND * iptw, 0))

# Sum the weighted outcomes and divide by the sum of weights for each group
ate_treat <- sum(TRC_data$weighted_outcome_treat) / 
  sum(TRC_data$iptw[TRC_data$TRCKNOW == 1])
ate_control <- sum(TRC_data$weighted_outcome_control) / 
  sum(TRC_data$iptw[TRC_data$TRCKNOW == 0])

ate_point_estimate <- ate_treat - ate_control

# Print the ATE point estimate
print("Average Treatment Effect Point Estimate")
ate_point_estimate

```
Interpretation 

- The negative sign of the estimate suggests that exposure to the TRC is
associated with a decrease in the level of agreement with the statement. 
In other words, those who are aware of the TRC are estimated to find it less 
difficult or challenging to understand the customs and ways of the opposite 
racial group, compared to those who are not aware of the TRC.
- The estimate has a relatively small magnitude, suggesting that while there is 
an association, the effect of the TRC exposure on racial attitudes or behavior
is modest. 
- In this scenario, we are assuming that the TRC exposure is conditionally 
ignorable given the covariates (i.e, no unmeasured confounders that affect both
the treatment assignment and the outcome), which means that the point estimate 
can be interpreted causally. Therefore, it suggests that the TRC exposure has a 
causal impact in reducing negative racial attitudes.

### Part d (10 points)

Using the bootstrap method (resampling individual rows of the data with replacement), obtain an estimate for the standard error of your IPTW estimator for the ATE. Compute a 95% confidence interval and interpret your findings. (You should report estimate, standard error, 95% CI lower, 95% CI upper, for interpretation, compare your results in Part C/D to your estimate from Part A and briefly discuss your findings.)


```{r}
# Set random seed
set.seed(123)
nBoot <- 1000 # Number of iterations
ate_boot <- rep(NA, nBoot) # Placeholder to store estimates

# For each iteration
for(boot in 1:nBoot){
  # Re-sample rows with replacement
  TRC_boot <- TRC_data[sample(1:nrow(TRC_data), nrow(TRC_data), replace=T),] 
  
  # Fit the propensity score model on the bootstrapped data
  pscore_model_boot <- glm(TRCKNOW ~ age + female + wealth + religiosity + 
                          ethsalience + rcblack + rcwhite + rccol + EDUC, data = 
                            TRC_boot, family = binomial(link = "logit"))
  
  # Save the propensities
  TRC_boot$e <- predict(pscore_model_boot, type = "response")
  
  # Calculate the weights
  TRC_boot$wt <- NA
  TRC_boot$wt[TRC_boot$TRCKNOW == 1] <- 1/TRC_boot$e[TRC_boot$TRCKNOW==1]
  TRC_boot$wt[TRC_boot$TRCKNOW == 0] <- 1/(1 - TRC_boot$e[TRC_boot$TRCKNOW==0])
  
  # Compute and store the ATE
  ate_boot[boot] <- mean(TRC_boot$wt * TRC_boot$RUSTAND * TRC_boot$TRCKNOW -
  TRC_boot$wt * TRC_boot$RUSTAND * (1-TRC_boot$TRCKNOW))}

# SD of the ate_boot to get our estimated SE 
sd(ate_boot)
```
```{r}
# Asymptotic 95% CI
c(ate_point_estimate - qnorm(.975)*sd(ate_boot),
ate_point_estimate + qnorm(.975)*sd(ate_boot))
```
```{r}
# Estimate
mean(ate_boot)
```
Interpretation: 

In part d, we implemented the bootstrap method to identify ATE using IPTW. The 
value of -0.1631028 represents the estimates difference in racial attitudes 
between those who were and were not exposed to the TRC. Similar to our 
interpretation in part a, the negative value of -0.1596519 indicates that 
TRC leads to a reduction in negative racial attitudes. To identify the 
significance of the effect, the confidence interval ranges from -0.25197297 to 
-0.0742325 and does not include 0, suggesting that the effect is indeed 
statistically significant at the 95% confidence level. 

Comparison:

The estimated ATE in part c implements a more directed approach without 
bootstrapping. As shown in the output, the point estimate from part c is 
close to the one in part d (-0.1631028). On the other hand, part a 
measures the effect of TRC exposure on racial attitudes using a direct linear
regression model. The coefficient for `TRCKNOW` was -0.2177, which is slightly 
larger in magnitude than the ATE estimates in parts c and d. The difference
might be due to the varying estimate methods. In other words, the coefficient in
part a represents the change in the outcome variable for each unit change in
the predictor, while part c and d (IPTW and bootstrapping) adjust for potential
confounders by weighting observations, which might lead to a more nuanced 
estimate of the treatment effect. 

Overall, all parts consistently show that knowledge about the TRC is associated
with a decrease in the difficulty of understanding customs and behaviors of the
opposite racial group, suggesting a positive impact of TRC exposure on racial
attitudes. The bootstrapped confidence intervals in part d provide a more solid
understanding of the estimate's variability and further validates the 
statistical significance of the effect. Although the effect sizes are fairly 
small, they are consistent and statistically significant across the different 
methods, reinforcing the importance of TRC exposure in influencing racial 
attitudes and that those with TRC exposure find it less difficult to understand 
the customs and ways of the opposite racial group.


## Question 2 (Total: 50 points)

Use the same data set as in Question 1.

### Part a (15 points)

Estimate the ATT of TRC exposure on respondents’ racial attitudes using the MatchIt approach. You can use the matchit function from MatchIt package in R. Implement the nearest neighbor matching algorithm and estimate the ATT. Report the 95% confidence interval of your estimate.

```{r}
library(MatchIt)
library(estimatr)

# Perform nearest neighbor matching
# For each treated unit, one nearest control unit will be selected as its match
matchit_model <- matchit(TRCKNOW ~ age + female + wealth + religiosity + 
                         ethsalience + rcblack + rcwhite + rccol + EDUC, 
                         data = TRC_data, method = "nearest", estimand = "ATT", 
                         distance = "glm")

# Estimate the ATT
matched_data <- match.data(matchit_model, weights = "new_weights")

att_model <- lm_robust(RUSTAND ~ TRCKNOW, data = matched_data, se_type = "HC1")

# Obtain the point estimate and robust standard error for ATT
att_estimate <- coef(summary(att_model))["TRCKNOW", "Estimate"]
att_se <- coef(summary(att_model))["TRCKNOW", "Std. Error"]

# Construct the 95% confidence interval
ci_lower <- att_estimate - qnorm(0.975) * att_se
ci_upper <- att_estimate + qnorm(0.975) * att_se

# Report the ATT and confidence interval
list(ATT = att_estimate, SE = att_se, CI_lower = ci_lower, CI_upper = ci_upper)

```
Interpretation: 

The value of -0.2140375 for the ATT estimate represents the average effect of 
TRC exposure on the racial attitudes of those respondents who were actually 
exposed to the TRC. A negative value indicates that, on average, exposure to the 
TRC is associated with a decrease in the level of agreement with the statement 
regarding difficulty in understanding customs and ways of the opposite racial 
group. In other words, those exposed to the TRC, on average, report a more
favorable or understanding attitude towards the customs and ways of the opposite 
racial group compared to what they would have reported had they not been 
exposed. The confidence interval ranges from -0.3054404 to -0.1226347 suggesting
that we can be 95% confident that the true ATT lies within this range. 
Considering that the entire interval is below zero strengthens the evidence that
TRC exposure has a statistically negative effect on the level of agreement with
the statement regarding the difficulty of understanding customs of the opposite
racial group. 

### Part b (15 points)

Estimate the ATT of TRC exposure on respondents’ racial attitudes using the MatchIt approach. You can use the matchit function from MatchIt package in R. Implement the exact matching algorithm and estimate the ATT. Report the 95% confidence interval of your estimate.

```{r}
library(MatchIt)
library(estimatr)

# Perform nearest neighbor matching
# For each treated unit, one nearest control unit will be selected as its match
matchit_model <- matchit(TRCKNOW ~ age + female + wealth + religiosity + 
                         ethsalience + rcblack + rcwhite + rccol + EDUC, 
                         data = TRC_data, method = "exact", estimand = "ATT", 
                         distance = "glm")

# Estimate the ATT
matched_data <- match.data(matchit_model, weights = "new_weights")

att_model <- lm_robust(RUSTAND ~ TRCKNOW, data = matched_data, se_type = "HC1")

# Obtain the point estimate and robust standard error for ATT
att_estimate <- coef(summary(att_model))["TRCKNOW", "Estimate"]
att_se <- coef(summary(att_model))["TRCKNOW", "Std. Error"]

# Construct the 95% confidence interval
ci_lower <- att_estimate - qnorm(0.975) * att_se
ci_upper <- att_estimate + qnorm(0.975) * att_se

# Report the ATT and confidence interval
list(ATT = att_estimate, SE = att_se, CI_lower = ci_lower, CI_upper = ci_upper)
```
Interpretation: 

Unlike the previous estimate with nearest neighbor matching, the ATT estimate
value is positive (0.1067194). This suggests that, on average, exposure to the 
TRC is associated with a slight increase in the level of agreement with the 
statement regarding difficulty in understanding customs and ways of the opposite 
racial group. However, given the context of the study, this might imply that 
those exposed to the TRC are slightly more likely to acknowledge differences or 
complexities in understanding the customs and ways of the opposite racial group.
While the standard error is relatively large in comparison to the ATT 
estimate (0.181528) indicating a higher level of uncertainty or variability in
the estimate, the wide confidence interval (ranging from -0.249069 to 0.4625077) 
which crosses zero suggests that the effect of TRC exposure on racial attitudes 
is actually not statistically significant at the 95% confidence level. Such a 
wide range in both positive and negative values indicates that the data is also 
consistent with a wide range of effects of TRC exposure on racial attitudes. 


### Part c (10 points)

Estimate the ATT of TRC exposure on respondents’ racial attitudes using the MatchIt approach. You can use the matchit function from MatchIt package in R. Implement the \textbf{coarsened exact matching} algorithm and estimate the ATT. Report the 95% confidence interval of your estimate.

```{r}
library(MatchIt)
library(estimatr)

# Perform nearest neighbor matching
# For each treated unit, one nearest control unit will be selected as its match
matchit_model <- matchit(TRCKNOW ~ age + female + wealth + religiosity + 
                         ethsalience + rcblack + rcwhite + rccol + EDUC, 
                         data = TRC_data, method = "cem", estimand = "ATT", 
                         distance = "glm")

# Estimate the ATT
matched_data <- match.data(matchit_model, weights = "new_weights")

att_model <- lm_robust(RUSTAND ~ TRCKNOW, data = matched_data, se_type = "HC1")

# Obtain the point estimate and robust standard error for ATT
att_estimate <- coef(summary(att_model))["TRCKNOW", "Estimate"]
att_se <- coef(summary(att_model))["TRCKNOW", "Std. Error"]

# Construct the 95% confidence interval
ci_lower <- att_estimate - qnorm(0.975) * att_se
ci_upper <- att_estimate + qnorm(0.975) * att_se

# Report the ATT and confidence interval
list(ATT = att_estimate, SE = att_se, CI_lower = ci_lower, CI_upper = ci_upper)

```
Interpretation: 

An ATT estimate of -0.1375487 indicates that, on average, exposure to the TRC is 
associated with a decrease in the level of agreement with the statement about 
difficulty understanding customs and ways of the opposite racial group among 
those respondents who were actually exposed to the TRC. A negative ATT value 
implies that TRC exposure leads to more positive or understanding attitudes 
towards the opposite racial group. The standard error value of 0.06311022 and a
95% confidence interval ranging from -0.2612424 to -0.01385493 suggests that the 
negative effect of TRC exposure on the stated measure of racial attitudes is 
statistically significant at the 95% confidence level. The interval also
indicates the range within which the true ATT is likely to fall.

### part d (10 points)

Compare and contrast the three different matching algorithms. Provide evidence and an argument about which one we should use.


Part a: Nearest Neighbor Matching 
- ATT Estimate: -0.2140375
- Standard Error: 0.04663497
- Confidence Interval: [-0.3054404, -0.1226347]

The nearest neighbor matching method often yields a good balance between both 
the treatment and control groups since it is flexible, and therefore does not 
require exact matches on covariates. This can be beneficial with continuous 
variables or large datasets, yet simultaneously can lead to sensitivity to the 
choice of distance metric. As a result, this could lead to inaccuracies in 
identifying the number of neighbors chosen, allowing for bad matches to take 
place if the nearest neighbor is not very close, where there are not enough 
close control units for each treated unit. 

Part b: Exact Matching
- ATT Estimate: 0.1067194
- Standard Error: 0.181528
- Confidence Interval: [-0.249069, 0.4625077]

The exact matching method ensures a perfect match on covariates, where 
treatment and control units that have identical values on all covariates are 
paired. As a result, this eliminates the possibility of model dependency in 
matching. However simultaneously, this also leads to a significant loss of data,
resulting in a very wide range of confidence intervals (as seen in our output), 
indicating lower precision in our estimates. 

Part c: Coarsened Exact Matching:
- ATT Estimate: -0.1375487
- Standard Error: 0.06311022
- Confidence Interval: [-0.2612424, -0.01385493]

The coarsened exact matching method provides a balance between exact matching
and nearest neighbor matching as it is a method that groups treatment and 
control units based on covariates and matches within these categories. It 
overall reduces the dimensionality of the covariate space, potentially leading
to better matches and a reduced loss of data. However in reality, it might still
lose some observations though generally less than exact matching. 


Considering the pros and cons of each method, the Nearest Neighbor Matching 
provides a balance between match quality and retaining enough data. The negative 
ATT estimate aligns with the expected impact of the TRC and has a relatively 
small confidence interval, indicating high precision. While Exact Matching 
ensures perfect matches, it resulted in a less precise estimate with a wider
confidence interval. The positive ATT estimate is counterintuitive given the 
context of TRC exposure, potentially reflecting the loss of valuable data. 
Coarsened Exact Matching seems to offer a middle ground. It provides a 
statistically significant negative ATT estimate, and has a moderately precise 
confidence interval. This method appears effective in balancing the data 
retention and match quality, making it a robust choice for estimating the ATT.

In conclusion, **Coarsened Exact Matching** might be the preferred approach for 
this analysis, considering its balance in data retention, match quality, and the 
intuitiveness and statistical significance of its results. However, by 
implementing a variety of matching algorithms, as done here, is necessary to 
identify the most appropriate choice and simultaneously ensure that the 
robustness of our approach is accurate in identifying causal inferences in our 
data. 

## BONUS ONLY: Question 3 (Total: Up to +12)
Question 3 is for bonus points. (See forthcoming lecture on Nov. 7th)

### part a (+4 points)
Using the regression method to predict potential outcomes for all individuals in the dataset and calculate the ATE with bootstrapped standard errors. Report and interpret your results. (Hint: Start by fitting the treatment and control model with subsets of the data.)
```{r}
library(tidyverse)
library(broom)
library(boot)
## Fit a model among TRCKNOW == 1 to get E[Y_i(1) | X]
# Subset the data into treated and control groups
treated_data <- TRC_data %>% filter(TRCKNOW == 1)
control_data <- TRC_data %>% filter(TRCKNOW == 0)

## Fit a model among TRCKNOW == 0 to get E[Y_i(0) | X]
model_treated <- lm(RUSTAND ~ age + female + wealth + religiosity + ethsalience 
                  + rcblack + rcwhite + rccol + EDUC, data = treated_data)
model_control <- lm(RUSTAND ~ age + female + wealth + religiosity + ethsalience 
                  + rcblack + rcwhite + rccol + EDUC, data = control_data)

## Predict the potential outcome under treatment for all units
TRC_data$predicted_treated <- predict(model_treated, newdata = TRC_data)
## Predict the potential outcome under control for all units
TRC_data$predicted_control <- predict(model_control, newdata = TRC_data)

## Average of the differences
ATE <- mean(TRC_data$predicted_treated - TRC_data$predicted_control)

# Bootstrap 
boot_func <- function(data, indices) {
  # Sample with replacement from the dataset
  bs_data <- data[indices, ]
  
  # Fit models and predict as above
  model_treated <- lm(RUSTAND ~ age + female + wealth + religiosity + 
                      ethsalience + rcblack + rcwhite + rccol + EDUC, data = 
                        bs_data[bs_data$TRCKNOW == 1, ])
  model_control <- lm(RUSTAND ~ age + female + wealth + religiosity + 
                      ethsalience + rcblack + rcwhite + rccol + EDUC, data = 
                        bs_data[bs_data$TRCKNOW == 0, ])
  
  bs_data$predicted_treated <- predict(model_treated, newdata = bs_data)
  bs_data$predicted_control <- predict(model_control, newdata = bs_data)
  
  # Return the ATE estimate
  mean(bs_data$predicted_treated - bs_data$predicted_control)
}

# Set the seed for reproducibility
set.seed(123)

# Perform bootstrap
boot_results <- boot(TRC_data, boot_func, R = 1000)

# Calculate bootstrapped standard error and confidence interval
boot_se <- sd(boot_results$t)
conf_int <- boot.ci(boot_results, type = "perc")$percent[4:5]

# Results
list(
  ATE = ATE,
  SE = boot_se,
  CI_lower = conf_int[1],
  CI_upper = conf_int[2]
)
```
Interpretation: 

An ATE estimate of -0.1743866 suggests that exposure to the TRC is associated 
with a decrease in the level of agreement with the statement about difficulty 
understanding customs and ways of the opposite racial group. In other words, 
TRC exposure leads, on average, to more favorable or understanding attitudes 
towards the opposite racial group. The standard error of 0.04403547 reflects the 
variability of the estimated effect size across different samples indicating 
higher precision in our estimate. As for the 95% confidence interval, the values
range from -0.2593992 to -0.085592, suggesting that we can be 95% confident that 
the true ATE lies within this range. The entire interval is below zero, further
reinforcing the finding that the effect of TRC exposure on racial attitudes is 
negative as well as statistically significant.

### part b (+4 points)
Using the regression method to predict potential outcomes for all individuals and calculate the ATT with bootstrapped standard errors. Report and interpret your results. 

```{r}
# Fit model to the control group
model_control <- lm_robust(RUSTAND ~ age + female + wealth + religiosity +
                           ethsalience + rcblack + rcwhite + rccol + EDUC, 
                           data = subset(TRC_data, TRC_data$TRCKNOW == 0))
model_treatment <- lm_robust(RUSTAND ~ age + female + wealth + religiosity +
                           ethsalience + rcblack + rcwhite + rccol + EDUC, 
                           data = subset(TRC_data, TRC_data$TRCKNOW == 1))
          

# Predict the potential outcome under treatment for all units
TRC_data$trc_treated <- predict(model_treatment, newdata = TRC_data)

# Predict the potential outcome under control for all units
TRC_data$trc_control <- predict(model_control, newdata = TRC_data)

# Average of the differences between observed and imputed potential outcomes
point_ATT <- mean(TRC_data$trc_treated[TRC_data$TRCKNOW==1] - 
                  TRC_data$trc_control[TRC_data$TRCKNOW==1])

# Bootstrap for SEs
set.seed(123)

nBoot <- 2000 # Number of iterations

boot_results_ATT <- rep(NA, 2000)

for (iter in 1:nBoot){
  # Resample w/ replacement
  treatment_boot <- TRC_data[sample(1:nrow(TRC_data), nrow(TRC_data), 
                                           replace=T),]
  
  # Fit a model to get E[Y_i(1) | X]
  treatment_model_boot <- lm_robust(RUSTAND ~ age + female + wealth + 
                                    religiosity + ethsalience + rcblack + 
                                    rcwhite + rccol + EDUC, data = 
                                    subset(treatment_boot, TRCKNOW == 1))
  # Fit a modelto get E[Y_i(0) | X]
  control_model_boot <- lm_robust(RUSTAND ~ age + female + wealth + 
                                    religiosity + ethsalience + rcblack + 
                                    rcwhite + rccol + EDUC, data = 
                                    subset(treatment_boot, TRCKNOW == 0))
  # Predict the potential outcome under treatment for all units
  treatment_boot$trc_treated_boot <- predict(treatment_model_boot, newdata = 
                                               treatment_boot)
  # Predict the potential outcome under control for all units
  treatment_boot$trc_control_boot <- predict(control_model_boot, newdata = 
                                              treatment_boot)
  # Store bootstrapped estimate
  boot_results_ATT[iter] <- mean(
    treatment_boot$trc_treated_boot[treatment_boot$TRCKNOW==1] -
    treatment_boot$trc_control_boot[treatment_boot$TRCKNOW==1])
}

# Standard error
standard_error = sd(boot_results_ATT)

# 95% confidence interval
ci = c(point_ATT - 1.96*sd(boot_results_ATT), point_ATT + 1.96*
         sd(boot_results_ATT))
# Results
list(
  ATT = point_ATT,
  SE = standard_error,
  CI_lower = ci[1],
  CI_upper = ci[2]
)
```
Interpretation: 

The ATT estimate has a value of -0.2033737, which suggests that those exposed to 
the TRC, on average, exhibit more positive or understanding racial attitudes
compared to what they would have exhibited had they not been exposed. The 
standard error value of 0.04641271 and our confidence interval, which ranges 
from -0.2943426 to -0.1124048, provides a range of values within which the true 
ATT is likely to lie with 95% confidence. Considering that the entire confidence 
interval is below zero and does not cross 0 further reinforces the evidence that 
the effect of TRC exposure on racial attitudes is negative and statistically
significant.

### part c (+4 points) 

Compare and contrast the ATE and ATT from the regression approach. 

The ATE estimate, as shown previously, is -0.1743866, while the ATT estimate is 
-0.2033737. Considering that the ATT is more negative than the ATE suggests that 
the treatment (TRC exposure) had a slightly stronger effect on those who were 
treated compared to the average effect on the entire population. This could 
indicate that those who were exposed to the TRC were more receptive or 
influenced by the exposure, leading to a more significant change in their 
racial attitudes.

In the context of the study and reality, the ATE is useful for understanding the 
overall impact of the intervention on the entire population. On the other hand,
the ATT is more relevant when considering the continuation or modification of 
the TRC, as it focuses on the impact on those who are actually receiving the 
treatment. In other words, the negative ATE value suggests that overall, TRC 
exposure is associated with more positive racial attitudes in the general 
population. The more negative ATT value additionally indicates that this effect
is even more prominent or stronger among those who actually received the 
treatment. In other words, this might indicate that the TRC was particularly 
effective for those who were exposed to the TRC were perhaps more open to 
change. 