---
title: "Problem Set 4"
author: "Ghina Al Shdaifat - gha2009 - 003 "
date: "Due Dec 13, 2023"
output:
  pdf_document: default
header-includes: 
  - \usepackage{tikz}
---

This homework must be turned in on Brightspace by Dec. 13 2023. It must be your own work, and your own work only -- you must not copy anyone's work, or allow anyone to copy yours. This extends to writing code. You may consult with others, but when you write up, you must do so alone.

Your homework submission must be written and submitted using Rmarkdown. No handwritten solutions
will be accepted. \textbf{No zip files will be accepted.} \textbf{Make sure we can read each line of code in the pdf document.} You should submit the following:

1. A compiled PDF file named yourNetID\_solutions.pdf containing your solutions to the problems.

2. A .Rmd file containing the code and text used to produce your compiled pdf named yourNetID\_solutions.Rmd.


Note that math can be typeset in Rmarkdown in the same way as Latex. Please make sure your answers are clearly structured in the Rmarkdown file:

1. Label each question part

2. Do not include written answers as code comments.

3. The code used to obtain the answer for each question part should accompany the written answer. Comment your code!


\newpage 

# Problem 1 (100 points)
Despite the heated political and media rhetoric, there are a few causal estimates of the effect of expanded health insurance on healthcare outcomes. One landmark study, the Oregon Health Insurance Experiment, covered new ground by utilizing a randomized control trial implemented by the state of Oregon. To allocate a limited number of eligible coverage slots for the state's Medicaid expansion, about 30,000 low-income, uninsured adults (out of about 90,000 wait-list applicants) were randomly selected by lottery to be allowed to apply for Medicaid coverage. Researchers collected observable measure of health (blood pressure, cholesterol, blood sugar levels, and depression), as well as hospital visitations and healthcare expenses for 6,387 selected adults and 5,842 not selected adults.

For this problem, we will use the OHIE.dta file.

* treatment - selected in the lottery to sign up for Medicaid (instrument)
* ohp_all_ever_admin - Ever enrolled in Medicaid after notification of lottery results (compliance) 
* tab2bp_hyper - Outcome: Binary indicator for elevated blood pressure (1 indicates a high blood pressure)
* tab2phqtot_high - Outcome: Binary indicator for depression
* tab4_catastrophic_exp_inp - Outcome: Indicator for catastrophic medical expenditure (1 if their total out-of-pocket medical expenses are larger than 30\% of their household income)
* tab5_needmet_med_inp - Outcome: Binary indicator of whether the participant feels that they received all needed medical care in past 12 months


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=F, messages=F, include=F}
library(estimatr)
library(tidyverse)
library(AER) # for waldtest
```

```{r}
# Load in the data
data <- haven::read_dta("OHIE.dta")
head(data)
```

\textbf{Hint:} This was an experiment with imperfect compliance. Instead of creating a "participated" or "complied" variable, simply use "treatment" as the instrument and "ohp_all_ever_admin" (enrollment in Medicaid) as the main independent variable of interest.

## Question A (25 points)

Estimate the intent-to-treat effects of being selected to sign up for Medicaid on each of the four outcomes (elevated blood pressure, depression, catastrophic medical expenditure, and whether respondents had their health care needs met). Provide 95\% confidence intervals for each estimate and interpret your results. (Use lm_robust)


```{r}
# Estimate the ITT on elevated blood pressure
itt_bp <- lm_robust(tab2bp_hyper ~ treatment, data = data) 
summary(itt_bp)

# Estimate the ITT on depression
itt_depression <- lm_robust(tab2phqtot_high ~ treatment, data = data) 
summary(itt_depression)

# Estimate the ITT on catastrophic expenditures
itt_catastrophic <- lm_robust(tab4_catastrophic_exp_inp ~ treatment, 
                              data = data) 
summary(itt_catastrophic)

# Estimate the ITT on "needs met"
itt_needs_met <- lm_robust(tab5_needmet_med_inp ~ treatment, data = data) 
summary(itt_needs_met)
```
Interpretation of results: 

The impact on elevated blood pressure reveals an estimate of -0.0016, where the 
confidence interval ranges from -0.01458 to 0.01138. The data suggests that the 
likelihood or severity of elevated blood pressure is not significantly altered 
by enrollment in Medicaid. This is evident from the confidence interval that 
crosses zero, implying no statistically significant change.

In regards to the effect of depression, the study indicates a significant 
reduction in depression among participants, with an estimate of -0.03493, with a
confidence interval that extends from -0.05102 to -0.01885. This negative 
estimate, coupled with a confidence interval that does not cross zero, strongly 
suggests that Medicaid enrollment may play a role in lessening depression 
symptoms or lowering the risk of depression.

There is a significant decrease in catastrophic medical expenditures, as shown 
by an estimate of -0.01527. The confidence interval lies between -0.02287 and 
-0.007665. This outcome implies that Medicaid enrollment potentially alleviates 
the financial burden associated with major medical expenses, as indicated by the 
negative estimate and a confidence interval that excludes zero.

The impact on meeting health care needs shows a positive effect, with an 
estimate of 0.03446 and a confidence interval that ranges from 0.01732 to 
0.0516. This positive estimate and its corresponding confidence interval suggest 
a significant improvement in the fulfillment of health care needs for those 
selected for Medicaid, indicating that Medicaid enrollees are more likely to 
have their health care requirements adequately addressed.

Overall, the findings in this specific scenario highlights that selection for 
Medicaid enrollment has a beneficial effect in reducing depression and 
catastrophic health expenditures, as well as in enhancing the satisfaction of 
health care needs. However, it does not appear to significantly affect elevated 
blood pressure levels. This analysis underscores the potential positive 
implications of Medicaid in specific health and financial domains.

## Question B (25 points)

Suppose that researchers actually wanted to estimate the effect of Medicaid enrollment (ohp_all_ever_admin) on each of the four outcomes. Suppose they first used a naive regression of each of the the outcomes on the indicator of Medicaid enrollment. Report a 95\% confidence interval for each of your estimates and interpret your results. Why might these be biased estimates for the causal effect of Medicaid enrollment?



```{r}

# Create a function to estimate Naive OLS effect and print the results
estimate_naive_regression <- function(outcome) {
model <- lm(as.formula(paste(outcome, "~ ohp_all_ever_admin")), data = data) 
summary_result <- summary(model)
ci <- confint(model, level = 0.95)
return(list(summary = summary_result, ci = ci))
}

# Estimate the Naive OLS effect on elevated blood pressure
naive_bp <- estimate_naive_regression("tab2bp_hyper") 
naive_bp

# Estimate the Naive OLS effect on depression
naive_depression <- estimate_naive_regression("tab2phqtot_high") 
naive_depression

# Estimate the Naive OLS effect on catastrophic expenditures
naive_expenses <- estimate_naive_regression("tab4_catastrophic_exp_inp") 
naive_expenses

# Estimate the Naive OLS on needs met
naive_care_needs_met <- estimate_naive_regression("tab5_needmet_med_inp") 
naive_care_needs_met

```
Interpretation of results: 

The confidence interval for the influence on elevated blood pressure ranges from 
-0.0324 to -0.00373. This outcome implies a marginal decrease in elevated blood 
pressure among those enrolled in Medicaid. The effect size is minor, and the 
range of the confidence interval is quite tight.

On the other hand, the impact on depression has a confidence interval that 
extends from 0.0316 to 0.0671. This suggests a possible rise in depression cases 
for Medicaid participants. The statistical significance of this effect is 
underlined by the confidence interval not covering zero.

The effect on catastrophic medical expenditures has a confidence interval that 
lies between -0.0191 and -0.00237. This negative estimate hints at a decrease in 
catastrophic medical costs due to Medicaid enrollment and is additionally 
statistically significant. 

With a confidence interval ranging from 0.0424 to 0.0802, the result points to 
an enhanced fulfillment of healthcare needs in Medicaid enrollees, a 
statistically significant finding.

Potential Bias:

The population enrolling in Medicaid may have distinct characteristics compared 
to non-enrollees. These characteristics, whether health-related, economic, or 
demographic, might influence both enrollment decisions and health outcomes, 
leading to skewed estimates. Key variables that influence both Medicaid
enrollment and health outcomes, yet not accounted for in the regression, can 
additionally lead to misleading results. It is challenging to pinpoint if 
Medicaid enrollment leads to specific health outcomes or if individuals with 
certain health conditions are more inclined to enroll in Medicaid. It is also
important to note that any errors in reporting or recording data related to 
Medicaid enrollment or health outcomes can also skew the findings.

In conclusion, while the initial regression approach sheds light on the 
association between Medicaid enrollment and various health outcomes, these 
findings could be skewed due to selection bias, missing important variables, 
the potential for a two-way relationship, and data inaccuracies. For a more 
accurate depiction of causality, methodologies like randomized trials or 
instrumental variable analysis would be more reliable.

## Question C (25 points)

Suppose we were to use assignment to treatment as an instrument for actually receiving Medicaid coverage.

Consider that not everyone who was selected to apply for Medicaid actually ended up applying and receiving coverage. Likewise, some applicants who were not selected to receive the treatment nevertheless were eventually covered. What were the compliance rates (the level of Medicaid enrollment) for subjects who were selected and subjects who were not selected? Use a ``first stage" regression to estimate the effect of being selected on Medicaid enrollment to estimate the compliance rates. Is the instrument of assignment-to-treatment a strong instrument for actual Medicaid enrollment?



```{r, warnings = F}
# First Stage OLS
first_stage_mod <- lm_robust(ohp_all_ever_admin ~ treatment, data = data)

# null model (compliance given an intercept only model)
null_mod<-lm_robust(ohp_all_ever_admin ~ 1, data=data)

# F - Stat for Instrument Strength (use waldtest)
instrument_strength <- waldtest(first_stage_mod, "treatment")
summary(first_stage_mod)
instrument_strength
```
Interpretation of results:

The coefficient for the treatment variable (being selected for Medicaid) is 
0.236, with a 95% confidence interval ranging from 0.225 to 0.248. The 
statistical significance of this coefficient is high with a p-value of less than
2e-16 (p value < 2e-16). The positive and statistically significant coefficient 
indicates that being selected to apply for Medicaid substantially increases the 
likelihood of actual enrollment. This is evident from the size of the 
coefficient and its confidence interval. A rather key measure of an instrument's 
strength is the F-statistic from the first stage regression. Here, the 
F-statistic is 1.61e+03, which is substantially larger than 10. Additionally, 
the relatively large R-squared value (0.0719) suggests that a significant 
portion of the variation in Medicaid enrollment is explained by the treatment 
assignment.

Based on these results, we can conclude that the instrument of 
assignment-to-treatment is indeed a strong instrument for actual Medicaid 
enrollment. The high F-statistic and the significant coefficient for the 
treatment variable in the first stage regression support this conclusion. In 
this scenario, the assignment to treatment serves as a strong instrument for 
Medicaid enrollment. This is confirmed by the substantial impact of the 
treatment assignment on enrollment, as evidenced by the significant coefficient 
and the robust F-statistic.

## Question D (25 points)

Now estimate the effect of Medicaid enrollment on each of the four outcomes using an instrumental variables strategy. Report a 95\% confidence interval for your estimates and interpret your results. Compare the estimates to those you obtained in Question B.



```{r}
run_iv_estimation <- function(outcome) { iv_model <- iv_robust(
formula = as.formula(paste(outcome, "~ ohp_all_ever_admin | treatment")),
data = data )
ci <- confint(iv_model, level = 0.95)
return(list(estimation = iv_model, ci = ci)) }

# Estimate the IV effect on elevated blood pressure (use iv_robust())
iv_bp <- run_iv_estimation("tab2bp_hyper") 
iv_bp

# Estimate the IV effect on depression
iv_depression <- run_iv_estimation("tab2phqtot_high") 
iv_depression

# Estimate the IV effect on catastrophic expenditures
iv_expenses <- run_iv_estimation("tab4_catastrophic_exp_inp") 
iv_expenses

# IV estimate on needs met
iv_care_needs_met <- run_iv_estimation("tab5_needmet_med_inp") 
iv_care_needs_met
```
Interpretation of results:

The IV estimate (-0.0063) of elevated blood pressure shows a negligible and 
statistically insignificant effect of Medicaid enrollment on elevated blood 
pressure, as indicated by the confidence interval including zero (-0.0574 to 
0.0448). Compared to the results from Question B, this finding is consistent in 
suggesting no significant impact on elevated blood pressure. On the other hand, 
the negative IV depression estimate (-0.138) has a confidence interval that does
not include zero (-0.202 to -0.0732), indicating a significant reduction in 
depression due to Medicaid enrollment. This effect is opposite in comparison
to Question B, suggesting potential bias in the naive OLS model. The IV 
results show a significant decrease in catastrophic medical expenditures, with 
an estimate value of -0.0604 and a confidence interval that ranges from -0.0906 
to -0.0301 being entirely below zero. This finding aligns with the results from 
Question B but shows a more substantial effect, indicating that Medicaid 
enrollment might have a stronger protective effect against catastrophic medical
expenses than previously estimated. The positive IV estimate (0.135), with a 
confidence interval (from 0.068 to 0.203) that does not include zero, suggests a 
significant improvement in meeting health care needs due to Medicaid enrollment. 
This result is in line with the findings from Question B but indicates a more 
substantial positive impact of Medicaid on meeting health care needs.

Overall, the IV estimates generally align with the direction of the effects 
observed in Question B but tend to show more pronounced effects. This could 
indicate that the naive estimates from Question B may have been biased, 
potentially underestimating the true effects of Medicaid enrollment on these 
outcomes. The instrumental variables approach likely provides a more accurate 
estimation of the causal effects of Medicaid enrollment by addressing potential
endogeneity and unobserved confounding factors.
